Impact ionization processes in a photodriven Mott insulator: influence of phononic dissipation

### 2022-11-05 resub_PRB/letters/reply_to_referees_current.txt

1. It is really important to discuss why DMFT approximation is
expected to be meaningful in quasi-2D case (1-layer Hubbard model
coupled to leads).

A1: As the Referee says, the system considered here is not strictly 2D. There is a coupling in the third direction due to the hybridization to the leads. Although these are noninteracting, they introduce a dissipation which suppresses 2D-coherence and makes nonlocal correlations less relevant.

On the other hand, it is true that antiferromagnetic (even short-range) correlations may play an important role in the spreading of photoexcited carriers, as discussed in Phys. Rev. Lett. 113, 076405
(2014). These nonlocal effects are not taken into account in our paper and certainly deserve future investigations.

As asked, we added in the conclusions a discussion on this point, with a reference to the paper mentioned above and to a comprehensive review of nonlocal correlations beyond DMFT [Rev. Mod. Phys. 90, 025003 (2018)].

### 2022-11-05 resub_PRB/Paper1.tex
Moved the comment on DMFT validity to the conclusions. Also slight modifications. Letter modified accordingly

For this goal we consider the simplified model shown in Fig.~\ref{fig:setup}, consisting of a Hubbard layer located between two leads. The leads have themselves a wide band, as expected for good conductors, which is effectively narrowed due to the weak coupling to intermediate layers which we take as simplified models for the contacts. Electrons on the Hubbard layer interact locally with acoustic phonons. An external periodic electric field induces a Floquet steady state with a current flowing from the lead with the lower to the lead with the higher chemical potential, thus transferring its energy. We address this periodic problem via nonequilibrium Floquet dynamical mean-field theory (F-DMFT)\resub{, which has been shown to give meaningful results in quasi-2D setups (see, e.g.~\cite{so.do.18}) as the one considered, in which the coupling in a third direction, due to the hybridization to the leads, introduces a dissipation which suppresses 2D-coherence and makes nonlocal correlations less relevant. However, it has been shown that antiferromagnetic (even short-range) correlations may play an important role in the spreading of photoexcited carriers~\cite{ec.we.14}. In this paper, we explicitly don't take into account these nonlocal effects~\cite{ro.ha.18}.}

@@

In order to study II in a Mott insulating layer, we adjust the bands to achieve the energy scheme shown in Fig.~\ref{fig:energy_setup}(a). We follow Ref.~\cite{so.do.18} and take $U=12$ and unless stated otherwise $E_0=2$. The fermion leads and the acoustic phonons have temperature $T=0.02$~\footnote{\resub{The value chosen for the temperature is small as compared with the other characteristic energy scales of the system, e.g. $\omega_{\text{ph}}$, $U$ and $W_{\text{b}}$.}}. We consider a particle-hole symmetric system with $\Gamma_{\text{l}}=\Gamma_{\text{r}}=0.37$, $t_{\text{l}}=t_{\text{r}}=1.7$, $\varepsilon_{\text{l}/\text{r}}=\mp 6$ and $v_{\text{l}}=v_{\text{r}}=v$. The parameters $t_{\text{l/r}}$, $\varepsilon_{\text{l/r}}$, and  $\Gamma_{\text{l/r}}$ are chosen such that the leads' DOS approximately overlaps with the Hubbard bands, i.e. they exhibit a \textquoteleft bandwidth\textquoteright~\footnote{Notice that the support of the leads' spectra are formally infinite. $W_{\text{b}}$ is the value at which the spectrum gets  suppressed.} $W_{\text{b}}\approx 8$ and are centered at the same position. The chemical potentials are set to $\mu_{\text{l}/\text{r}} =\mp 1$, so that a current from left to right is produced by taking energy from the driving. Due to the hybridization with the leads, the local DOS of the Hubbard layer only features a pseudogap
 $\Delta_{\text{pg}} \approx 4$. All simulations are carried out with the dimensionless factor $\alpha\equiv t_{\text{resc}}E_0/\Omega^{2}<0.5$ for which the FDSA is justified~\footnote{In Ref.~\cite{so.do.18} the factor is defined as $\alpha=t_{c}E_0/\Omega^{2}$ and $t_{\text{c}}=1$. Setting $t_{\text{resc}}=1$, we define for consistency $\alpha$ as in the main text.}. Whenever the electron-phonon interaction is included, we take $g=0.8$ and unless stated otherwise $\omega_{\text{ph}}=0.1$.

@@

Experimentally, the electric field amplitude considered in this paper is several orders of magnitude larger than that of the sunlight, therefore our results are more relevant for photoexcitations produced by intense laser pulses (see also the discussion in~\cite{mu.we.18}) rather than for true photovoltaic systems. \resub{We did not consider smaller electric field amplitudes due to the slow F-DMFT convergence for such a choice of parameters. This also occurs for weaker hybridization strengths.}

On the other hand, one may try to qualitatively extrapolate the discussion at the end of Sec.~\ref{sec:E0_2_only_electrons} in the other direction to smaller electric field amplitudes, for which a smaller hybridization would be necessary in order to sustain II. In a more realistic setup modelling oxide heterostructures~\cite{as.bl.13,pe.be.19}, a regime of effectively small hybridization would be achieved by considering several layers in which photoexcited carriers, separated by an electric field gradient, have more time to induce II processes before escaping into the leads. To address the occurrence of II in such a setup, further extensions such as considering multiple orbitals and impurity scattering should be taken into account. \resub{Finally, a more realistic description of the effects of the electron-phonon interaction on electronic band structure should take into account the crystal momentum dependence of $\kel{\mat{\Sigma}}_{\text{e-ph}}$~\cite{gius.17}.} These would be interesting for future studies.

>>>>

For this goal we consider the simplified model shown in Fig.~\ref{fig:setup}, consisting of a Hubbard layer located between two leads. The leads have themselves a wide band, as expected for good conductors, which is effectively narrowed due to the weak coupling to intermediate layers which we take as simplified models for the contacts. Electrons on the Hubbard layer interact locally with acoustic phonons. An external periodic electric field induces a Floquet steady state with a current flowing from the lead with the lower to the lead with the higher chemical potential, thus transferring its energy. We address this periodic problem via nonequilibrium Floquet dynamical mean-field theory (F-DMFT).

@@

In order to study II in a Mott insulating layer, we adjust the bands to achieve the energy scheme shown in Fig.~\ref{fig:energy_setup}(a). We follow Ref.~\cite{so.do.18} and take $U=12$ and unless stated otherwise $E_0=2$. The fermion leads and the acoustic phonons have temperature $T=0.02$~\footnote{\resub{The value chosen for the temperature is much smaller as compared with the other characteristic energy scales of the system, e.g. $\omega_{\text{ph}}$, $U$ and $W_{\text{b}}$.}}. We consider a particle-hole symmetric system with $\Gamma_{\text{l}}=\Gamma_{\text{r}}=0.37$, $t_{\text{l}}=t_{\text{r}}=1.7$, $\varepsilon_{\text{l}/\text{r}}=\mp 6$ and $v_{\text{l}}=v_{\text{r}}=v$. The parameters $t_{\text{l/r}}$, $\varepsilon_{\text{l/r}}$, and  $\Gamma_{\text{l/r}}$ are chosen such that the leads' DOS approximately overlaps with the Hubbard bands, i.e. they exhibit a \textquoteleft bandwidth\textquoteright~\footnote{Notice that the support of the leads' spectra are formally infinite. $W_{\text{b}}$ is the value at which the spectrum gets  suppressed.} $W_{\text{b}}\approx 8$ and are centered at the same position. The chemical potentials are set to $\mu_{\text{l}/\text{r}} =\mp 1$, so that a current from left to right is produced by taking energy from the driving. Due to the hybridization with the leads, the local DOS of the Hubbard layer only features a pseudogap
 $\Delta_{\text{pg}} \approx 4$. All simulations are carried out with the dimensionless factor $\alpha\equiv t_{\text{resc}}E_0/\Omega^{2}<0.5$ for which the FDSA is justified~\footnote{In Ref.~\cite{so.do.18} the factor is defined as $\alpha=t_{c}E_0/\Omega^{2}$ and $t_{\text{c}}=1$. Setting $t_{\text{resc}}=1$, we define for consistency $\alpha$ as in the main text.}. Whenever the electron-phonon interaction is included, we take $g=0.8$ and unless stated otherwise $\omega_{\text{ph}}=0.1$.

@@

\resub{One should comment on the validity of DMFT, which should be accurate for large dimensions only, for the present quasi 2-D setup. As a matter of fact, the hybridization to the leads introduces a dissipation that suppresses 2D-coherence and thus makes nonlocal correlations beyond DMFT less relevant.
On the other hand, it has been shown that antiferromagnetic (even short-range) correlations may play an important role in the spreading of photoexcited carriers~\cite{ec.we.14}. 
Thaking into account the effect of such nonlocal correlations (see also ~\cite{ro.ha.18}) could be an interesting extension, although beyond the scope of the present research.}

Experimentally, the electric field amplitude considered in this paper is several orders of magnitude larger than that of the sunlight, therefore our results are more relevant for photoexcitations produced by intense laser pulses (see also the discussion in~\cite{mu.we.18}) rather than for true photovoltaic systems. \resub{We did not consider smaller electric field amplitudes due to the slow F-DMFT convergence for such a choice of parameters. This is also the case for weaker hybridization strengths.}

On the other hand, one may try to qualitatively extrapolate the discussion at the end of Sec.~\ref{sec:E0_2_only_electrons} in the other direction to smaller electric field amplitudes, for which a smaller hybridization would be necessary in order to sustain II. In a more realistic setup modelling oxide heterostructures~\cite{as.bl.13,pe.be.19}, a regime of effectively small hybridization would be achieved by considering several layers in which photoexcited carriers, separated by an electric field gradient, have more time to induce II processes before escaping into the leads. To address the occurrence of II in such a setup, further extensions such as considering multiple orbitals and impurity scattering should be taken into account. This aspect could be interesting for future studies.    
 \resub{Finally, a more realistic description of the effects of the electron-phonon interaction on electronic band structure should take into account the crystal momentum dependence of $\kel{\mat{\Sigma}}_{\text{e-ph}}$~\cite{gius.17}.}
